---
title: "FinalReport"
author: "Alexander Danilychev Jr (23574348)"
date: "December 10, 2015"
output: html_document
---

Placeholder text woo

__TaxRevenue__: Tax Revenue as % of GDP

__Political__, __Social__, __Internet__, __Military__ can all be found in the censorship README

## Sourcing and loading our dataframes

```{r}
if (!("corrplot" %in% rownames(installed.packages()))){
  install.packages("corrplot",repos="http://cran.rstudio.com/")
}
library(corrplot)
library(ggplot2)
library(reshape2)
source("graphHelper.R")
resetPar <- function() {
    dev.new()
    op <- par(no.readonly = TRUE)
    dev.off()
    op
}
censorship = read.csv("../cleandata/censorship.csv")
foreignAidDonors = read.csv("../cleandata/netForeignAid.csv")
foreignAidRecipients = read.csv("../cleandata/foreignAidRecipients.csv")
```

####Countries in Censorship dataset:

```{r}
print(censorship$Country)
```

####Countries in ForeignAid Donors dataset:

```{r}
print(foreignAidDonors$Country)
```

####Countries in ForeignAid Recipients dataset:

```{r}
print(foreignAidRecipients$Country)
```

##Aid Graphs (by itself)

####Spread of Foreign Aid Donors

```{r}
graphFullAid(foreignAidDonors)
```

####Graph of country and amount of aid given (just upper 3rd quartile)

```{r}
graphThirdQuartileAid(foreignAidDonors)
```

####Foreign Aid Recipient Full Spread

```{r}
fullForeignAidSpread(foreignAidRecipients)
```

####Top 10 ForeignAidRecipients
```{r}
tenHighestReceived(foreignAidRecipients)
```

####Frequencies of each kind of censorship

```{r}

highestCensorship(censorship, category = "Social")
highestCensorship(censorship, category = "Political")
highestCensorship(censorship, category = "Internet")
highestCensorship(censorship, category = "Military")
#Total censorship created by summing up the four categories as integers 1-4
highestCensorship(censorship, category = "TotalCensorshipScore")
```

```{r}
highestCensorship(censorship, category = "Social", "GDP_per_cap")
highestCensorship(censorship, category = "Political", "GDP_per_cap")
highestCensorship(censorship, category = "Internet", "GDP_per_cap")
highestCensorship(censorship, category = "Military", "GDP_per_cap")
```

##Checking correlations

###Lets try checking correlation including tax (around 20 tax values out of 72 are missing, so this is less countries)

```{r}
plotCorrMatrix(censorship, rmTax = FALSE, method = "ellipse")

plotCorrMatrix(censorship, rmTax = FALSE, method = "shade")

```

###Now lets try checking correlation without tax (more data included this way)

```{r}
plotCorrMatrix(censorship, rmTax = TRUE, method = "ellipse")

plotCorrMatrix(censorship, rmTax = TRUE, method = "shade")

```
